---
title: "lab4"
author: "Nathan Haas", "Alex Beagle"
date: "10/17/2019"
output: html_document
---

Owner Github username: nahaas97
Parter Github username: abeagle15
Repo: https://github.com/nahaas97/ds202_lab4

```{r}
library(readxl)
defense <- read_excel('cyclonesFootball2018.xlsx', sheet='Defensive')
str(defense)
offense <- read_excel('cyclonesFootball2018.xlsx', sheet='Offensive')
str(offense)
bio <- read_excel('cyclonesFootball2018.xlsx', sheet='Biography')
str(bio)

```
Cleaning the data
```{r}
library(dplyr)
#1.
defense$Name = factor(defense$Name)
offense$Name = factor(offense$Name)
bio$Name = factor(bio$Name)
offense$Opponent_Opponent = factor(offense$Opponent_Opponent)
defense$Opponent_Opponent = factor(defense$Opponent_Opponent)

#2.
defense$Tackles_Solo = as.numeric(defense$Tackles_Solo)
defense$Tackles_ASST = as.numeric(defense$Tackles_ASST)
defense$Tackles_TFL = as.numeric(defense$Tackles_TFL)
defense$Tackles_Sack = as.numeric(defense$Tackles_Sack)
defense$Turnover_FF = as.numeric(defense$Turnover_FF)
defense$Turnover_FR = as.numeric(defense$Turnover_FR)
defense$Turnover_INT = as.numeric(defense$Turnover_INT)
defense$Pass_QBH = as.numeric(defense$Pass_QBH)
defense$Pass_PB = as.numeric(defense$Pass_PB)

offense$Receiving_REC = as.numeric(offense$Receiving_REC)
offense$Receiving_YDS = as.numeric(offense$Receiving_YDS)
offense$Receiving_TD = as.numeric(offense$Receiving_TD)
offense$Rushing_ATT = as.numeric(offense$Rushing_ATT)
offense$Rushing_YDS = as.numeric(offense$Rushing_YDS)
offense$Rushing_TD = as.numeric(offense$Rushing_TD)
offense$`Passing_CMP-ATT` = as.numeric(offense$`Passing_CMP-ATT`)
offense$Passing_YDS = as.numeric(offense$Passing_YDS)
offense$Passing_TD = as.numeric(offense$Passing_TD)
offense$Passing_INT = as.numeric(offense$Passing_INT)

bio$Weight = as.numeric(bio$Weight)

library(tidyr)
#3.
bio = separate(bio, col="Height", into= c("Height_in_feet","Remaining_height_inches"),sep="-")
bio$Height_in_feet = as.numeric(bio$Height_in_feet)
bio$Remaining_height_inches = as.numeric(bio$Remaining_height_inches)
bio$Height_in_feet = bio$Height_in_feet * 12
bio$Height_in_inches = bio$Height_in_feet + bio$Remaining_height_inches
bio$Height_in_feet = NULL
bio$Remaining_height_inches = NULL
head(bio)

#4.
defClean = defense
str(defClean)
offClean = offense
str(offClean)
bioClean = bio
str(bioClean)
```


Tidying the Data (A)
```{r}
#1.
defClean = gather(defClean, key=Statistic, value=Count, 3:11)
library(ggplot2)

#2.
ggplot(defClean, aes(x=Count)) + geom_histogram(binwidth = 1) + facet_wrap(~Statistic) + labs(title="Count per defensive statistic", x="Number per player", y="Count")

```

The rarest defensive skills are interceptions (Turnover_INT) and fumble recoveries (Turnover_FR). While most of the skills are relatively rare, the two above are the rarest.


```{r}
#3.
library(ggplot2)
betterdefense = filter(defClean, Statistic=="Tackles_Solo") %>%
spread(key= Opponent_Opponent, value = Count)
ggplot(betterdefense, aes(x=`Kansas State`, y=`West Virginia`)) + geom_point() + labs(title = "Solo tackles against Kansas State and West Virginia", x="Solo Tackles against Kansas State", y="Solo Tackles against West Virginia")

```

According to the graph above, we can see that we had more solo tackles against West Virginia, meaning that we had a better defense against them compared to Kansas State.

Tidying (part B)

```{r}
#1.
bioClean = separate(bioClean, col="Hometown", into= c("City","State"),sep=",")
head(bioClean$City)
head(bioClean$State)

#2.
table(bioClean$State)
```

Joining Data Frames

1. 
```{r}
offensive_bio = left_join(offClean, bioClean, by= "Name")
head(offensive_bio)

table(unique(offensive_bio$City))
table(offensive_bio$Name)
```

2.
```{r}
full_offensive_bio = full_join(offClean, bioClean, by= "Name")

ggplot(full_offensive_bio, aes(Weight, Receiving_YDS)) + geom_point() + labs(title="Number of Receiving Yards by Weight", x="Weight (lbs)",y="Receiving Yards")
```

Between the weights of 175 lbs - 225 lbs., the number of receiving yards typically decreases as the weight of a player increases. There is an abnormality at the weight of 225, then the trend decreases again, generally.


3.
Is there a trend between the height of a player and the number of receptions?
```{r}
ggplot(full_offensive_bio, aes(Height_in_inches, Receiving_REC)) + geom_point(alpha=1/5) + labs(title="Number of Receptions by Height", x="Height (in)", y="Number of Receptions")
```

From the data we have gathered, there appears to be no general trend between height and the number of receptions. This may be because there are star players on the team at set heights, and there are a limited number of players in the dataset.

